% main.tex — NAI journal submission
% NOTE: adapt to IOS Press iosart class before submission.
% This draft uses standard article class for development.
\documentclass[lettersize,journal]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\geometry{left=3cm,right=3cm,top=2.8cm,bottom=2.8cm}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage[all]{xy}
\usepackage{hyperref}
\usepackage{url}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}

% ── Theorem environments ──────────────────────────────────────────────────────
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

% ── Notation macros ───────────────────────────────────────────────────────────
\newcommand{\tnorm}{\otimes}      % conjunction (t-norm)
\newcommand{\resid}{\Rightarrow}  % implication (residuum)
\newcommand{\disj}{\oplus}        % disjunction
\newcommand{\LNN}{\L{}NN}
\newcommand{\Luk}{\L{}ukasiewicz}

% ─────────────────────────────────────────────────────────────────────────────
\begin{document}

\title{\L{}ukasiewicz Neural Networks Extended: Residual Architectures and
       Crystallization Strategies for Interpretable Rule Extraction}

\author{Carlos Leandro\\
  \textit{Departamento de Matem\'atica,
  Instituto Superior de Engenharia de Lisboa (ISEL)},\\
  \textit{Instituto Polit\'ecnico de Lisboa, Portugal}\\
  \texttt{carlos.leandro@solidreturn.dev}}

\date{}
\maketitle

% ── Abstract ─────────────────────────────────────────────────────────────────
\begin{abstract}
A feed-forward neural network whose weights are restricted to $\{-1,0,+1\}$
and whose activation is the truncated identity implements the connectives of
\L{}ukasiewicz many-valued logic exactly, neuron by neuron.
Once weights are driven to integers --- a process called \emph{crystallization}
--- the network is not a black-box but a logical formula readable directly from
its topology.
Two obstacles have constrained this approach: the original
Levenberg--Marquardt (LM) training fails on high-dimensional problems, and the
symbolic guarantees held only for shallow architectures.

This paper resolves both.
We prove that \emph{residual skip connections} extend \L{}ukasiewicz networks
to arbitrary depth while preserving the symbolic correspondence by construction:
merge neurons automatically satisfy the neuron-classification criterion
regardless of what the inner layers compute.
We then compare three crystallization strategies --- corrected LM,
straight-through estimation (STE), and proximal regularization --- in both flat
and residual variants across six UCI benchmarks (10 independent trials each).
STE with residual connections achieves the highest mean $F_1$ on four of six
benchmarks and significantly outperforms its flat counterpart on structured
datasets.
LM with residual connections recovers exact logical formulas on structured
problems.
Proximal regularization collapses to majority-class predictions on five of six
datasets regardless of architecture, a structural failure mode that is
theoretically characterized.
A comparison with contemporaneous differentiable logic gate networks identifies
the complementary strengths of the two approaches.

\noindent\textbf{Keywords:} neuro-symbolic AI; \L{}ukasiewicz logic;
interpretable machine learning; symbolic rule extraction; residual networks;
weight crystallization; straight-through estimation.
\end{abstract}

% ── 1. Introduction ──────────────────────────────────────────────────────────
\section{Introduction}
\label{sec:intro}

The tension between predictive power and interpretability in machine learning
has deepened as models have grown larger.
A neural network that achieves human-level performance on a medical diagnosis
task may be clinically useless if its reasoning cannot be audited.
Rudin~\cite{rudin2019} argues that in high-stakes domains the correct response
is not to add a post-hoc explanation layer on top of a black-box model, but to
require that the model itself be interpretable by construction.
Post-hoc explanation methods provide approximations --- and an approximation
to a decision can be systematically wrong in ways that are invisible to the
explainer~\cite{guidotti2018}.

Symbolic rule extraction from neural networks has pursued this goal since the
1980s~\cite{gallant1988,towell1993,towell1994}.
The dominant paradigm trains a network, then searches for a logical rule that
approximates its input-output behaviour.
The approximation step is unavoidable in classical two-valued logic: a neuron
computing a weighted sum is not, in general, a Boolean connective.
\L{}ukasiewicz logic removes this obstacle.
Its connectives --- conjunction, implication, negation, and disjunction ---
are all piecewise linear functions on $[0,1]$, the same domain that
truncated-identity neurons compute over.
Castro and Trillas~\cite{castro1998} observed that this structural coincidence
is not accidental: a neuron with integer weights $\{-1,0,1\}$ and truncated
identity activation \emph{is} a \L{}ukasiewicz connective, exactly, without
approximation.
Leandro~\cite{leandro2009} turned this observation into a complete
methodology: train a feed-forward network so that its weights crystallize to
integers, then read the resulting formula directly from the network topology.

The programme is mathematically clean and practically motivated, but two
limitations have constrained its reach.
First, the Levenberg--Marquardt (LM) training algorithm used in the original
work crystallizes weights reliably on small, low-dimensional datasets, but
suffers from a numerical instability when weights are near zero at
initialisation --- a regime that \L{}ukasiewicz networks enter naturally ---
causing divergence before convergence.\footnote{The instability arises from the
Marquardt-form damping $\mu\,\mathrm{diag}(J^\top J)$: when weights are small
the diagonal entries of $J^\top J$ vanish, making the system ill-conditioned.
Section~\ref{ssec:lm} derives the corrected Levenberg form and quantifies
the magnitude of the divergence.}
We identify and correct this instability; the correction restores LM's
effectiveness on problems where it previously failed.
Second, the symbolic correspondence was proven only for networks with a single
hidden layer of classifiable neurons~\cite{castro1998,leandro2009}; it does not
extend trivially to deeper architectures, because training cannot guarantee
that every neuron in every additional layer satisfies the classification
criterion.

This paper extends the \L{}ukasiewicz neural network (\LNN{}) framework along
both dimensions.
On the theoretical side, we show that \emph{residual connections}
--- the skip-connection mechanism introduced by He~et~al.~\cite{he2016} ---
can be added to \LNN{}s while preserving the symbolic correspondence by
construction.
The key insight is architectural: the neuron that combines the skip path and
the inner-layer path --- the \emph{merge neuron}, which receives exactly two
inputs each with weight $+1$ --- automatically satisfies the
neuron-classification criterion regardless of what the inner layers compute.
On the empirical side, we compare three crystallization strategies (corrected
LM, straight-through estimation~\cite{bengio2013}, and proximal regularization)
in both flat and residual architectures, characterizing the conditions under
which each strategy succeeds or fails.

\paragraph{Contributions.}
\begin{enumerate}
  \item \textbf{Residual \L{}ukasiewicz networks (theory).}
    We prove that merge neurons in a \L{}ukasiewicz residual block satisfy
    the neuron-classification property (Proposition~\ref{prop:classif}) by
    construction, extending the symbolic framework to arbitrary depth.
  \item \textbf{Crystallization analysis.}
    We identify and correct a numerical instability in the original LM
    training, and characterize three crystallization strategies (LM-corrected,
    STE, proximal) with their theoretical guarantees and failure modes.
  \item \textbf{Empirical evaluation.}
    Five training configurations across six UCI benchmarks (10 independent
    trials each) show that STE with residual connections significantly
    outperforms flat STE on structured datasets and achieves the best
    aggregate discriminative quality overall.
  \item \textbf{Formula extraction in practice.}
    We demonstrate extraction of exact logical formulas from trained networks,
    including a clinically interpretable rule for cardiac disease diagnosis.
\end{enumerate}

The remainder of the paper is organized as follows.
Section~\ref{sec:background} reviews \L{}ukasiewicz logic and the original
\LNN{} framework.
Section~\ref{sec:residual} introduces residual \LNN{}s and proves the
merge-neuron theorem.
Section~\ref{sec:crystal} analyses the three crystallization strategies.
Section~\ref{sec:experiments} presents the experimental evaluation.
Section~\ref{sec:related} positions the work relative to the literature.
Sections~\ref{sec:discussion} and~\ref{sec:conclusion} discuss implications
and conclude.

% ── 2. Background ─────────────────────────────────────────────────────────────
\section{Background}
\label{sec:background}

\subsection{\L{}ukasiewicz many-valued logic}
\label{ssec:logic}

In classical propositional logic, truth values are drawn from $\{0,1\}$.
\L{}ukasiewicz logic generalizes this to the real unit interval $[0,1]$,
interpreting $0$ as complete falsity and $1$ as complete truth, with
intermediate values representing degrees of certainty~\cite{hajek1998}.
The four fundamental connectives are defined as follows.
\begin{definition}[\L{}ukasiewicz connectives]
For $x, y \in [0,1]$:
\begin{align}
  x \tnorm y  &= \max(0,\, x + y - 1),   \label{eq:tnorm}\\
  x \resid y  &= \min(1,\, 1 - x + y),   \label{eq:resid}\\
  \neg x       &= 1 - x,                  \label{eq:neg}\\
  x \disj y   &= \min(1,\, x + y).       \label{eq:disj}
\end{align}
\end{definition}
The operations $\tnorm$ and $\disj$ recover Boolean conjunction and disjunction
when restricted to $\{0,1\}$; they are, however, genuinely many-valued on $(0,1)$.
For instance, $\tfrac{1}{2} \tnorm \tfrac{1}{2} = 0$: two half-truths do not
compose into a truth under conjunction unless their sum exceeds $1$.
Note that $x \disj y = \neg(\neg x \tnorm \neg y)$, so disjunction is
De~Morgan dual to conjunction, as in classical logic.
The residuum $\resid$ is the natural implication associated with the t-norm
$\tnorm$~\cite{hajek1998}.

A \emph{McNaughton function}~\cite{mcnaughton1951} is a continuous, piecewise
linear map $f\colon [0,1]^n \to [0,1]$ with integer coefficients in each linear
piece.
McNaughton's theorem states that $f$ is a McNaughton function if and only if
it is expressible as a formula in infinitely-valued \L{}ukasiewicz
logic~\cite{mcnaughton1951,mundici1986}.
All four connectives in Eqs.~\eqref{eq:tnorm}--\eqref{eq:disj} are
piecewise linear with integer coefficients, confirming that the logic is
exactly characterised by this function class.

For any integer $n \geq 1$, the finite set
$S_n = \{0, 1/n, 2/n, \ldots, 1\}$ is closed under all four operations.
These finite subdomains allow truth-table-style evaluation and make the
semantics computationally tractable.

\subsection{\L{}ukasiewicz neural networks}
\label{ssec:lnn}

A neuron with $m$ inputs $x_1, \ldots, x_m \in [0,1]$, weights
$w_1, \ldots, w_m \in \mathbb{R}$, bias $b \in \mathbb{R}$, and
\emph{truncated identity activation}
\begin{equation}
  \psi(x) = \min\!\bigl(1,\, \max(0,\, x)\bigr),
  \label{eq:psi}
\end{equation}
produces output $z = \psi\!\bigl(\sum_{i=1}^m w_i x_i + b\bigr)$.
We extend the notation to include the bias explicitly:
\begin{equation}
  \psi_b(z) \;=\; \psi(z + b) \;=\; \min\!\bigl(1,\,\max(0,\, z + b)\bigr).
  \label{eq:psib}
\end{equation}
This activation is piecewise linear with integer coefficients, matching the
structure of McNaughton functions.

The following proposition, established by Castro and
Trillas~\cite{castro1998} and extended by Leandro~\cite{leandro2009},
is the central result connecting neural computation to \L{}ukasiewicz logic.

\begin{proposition}[Neuron classification~\cite{castro1998,leandro2009}]
\label{prop:classif}
Let $\alpha = \psi_b(-x_1, \ldots, -x_n,\, x_{n+1}, \ldots, x_m)$
be a neuron with $n$ weights equal to $-1$ and $p = m - n$ weights equal to
$+1$.
\begin{enumerate}
  \item If $b = -p + 1$, then $\alpha$ computes the \emph{conjunction}
    $\neg x_1 \tnorm \cdots \tnorm \neg x_n \tnorm x_{n+1}
     \tnorm \cdots \tnorm x_m$.
  \item If $b = n$, then $\alpha$ computes the \emph{disjunction}
    $\neg x_1 \disj \cdots \disj \neg x_n \disj x_{n+1}
     \disj \cdots \disj x_m$.
\end{enumerate}
\end{proposition}

\begin{proof}[Proof sketch]
For case 1 with $n=0$: $\psi_{-p+1}(x_1+\cdots+x_m)
= \max(0, x_1+\cdots+x_m-(p-1)) = x_1 \tnorm \cdots \tnorm x_m$,
which follows by induction using $x \tnorm y = \max(0,x+y-1)$.
For $n > 0$, substitute $\neg x_i = 1-x_i$ for each negated input:
the weighted sum becomes
$\sum_{j=n+1}^m x_j - \sum_{i=1}^n x_i + n + b$.
Setting $b = -p + 1 = -(m-n) + 1$ yields
$\sum_{j=n+1}^m x_j + \sum_{i=1}^n(1-x_i) - 1$,
and $\psi$ applied to this recovers the $m$-fold conjunction.
Case 2 is analogous using $b = n$ and the disjunction definition.
\end{proof}

A network in which \emph{every} neuron satisfies Proposition~\ref{prop:classif}
is a \emph{\L{}ukasiewicz neural network} (\LNN{}).
In a \LNN{}, the output can be read as a \L{}ukasiewicz formula by composing
the sub-formulas computed at each neuron, propagating bottom-up through the
layers.
Conversely, any \L{}ukasiewicz formula can be encoded as a \LNN{} by mapping
each connective to a neuron with the corresponding bias and weights
$\{-1, 0, +1\}$~\cite{amato2002}.
This bidirectional correspondence --- formula $\leftrightarrow$ network ---
is the foundation of the symbolic extraction methodology.

\subsection{Rewriting algebra and symbolic extraction}
\label{ssec:rewriting}

Proposition~\ref{prop:classif} covers neurons with at most two distinct weight
values ($-1$ and $+1$).
Real networks, after training, may have neurons with more complex weight
configurations that do not directly satisfy the proposition.
The \emph{rewriting algebra} of~\cite{leandro2009} handles this case.

\paragraph{Rule R.}
Rule~R peels one input off the neuron by introducing a binary sub-neuron with
connection weight $+1$; the original bias is distributed as $b = b_0 + b_1$.
The split is valid when (i)~the original bias is distributed as $b = b_0 + b_1$,
where $b_0$ is assigned to the left sub-neuron and $b_1$ to the right
sub-neuron, and (ii)~neither sub-neuron produces a constant output
(i.e.\ neither $\psi_{b_0}$ nor $\psi_{b_1}$ is identically $0$ or $1$
on the relevant domain).
Applying Rule~R exhaustively to all valid bias distributions $(b_0, b_1)$
generates a finite set of candidate two-neuron networks.

\paragraph{$\lambda$-similarity.}
The proximity between two networks over a finite subdomain $S_n$ is measured by
\begin{equation}
  \lambda = \exp\bigl(-\mathrm{MAE}(f(X),\, y)\bigr),
  \label{eq:lambda}
\end{equation}
where $\mathrm{MAE}$ is the mean absolute error over all input vectors
$X \in S_n^m$.
When $\lambda = 1$, the two functions are identical on $S_n$.
Since outputs lie in $[0,1]$, $\mathrm{MAE} \leq 1$ and hence
$\lambda \geq e^{-1} \approx 0.368$; smaller values of $\lambda$ indicate
greater divergence between the two functions.
A neuron is \emph{representable} in the sense of Proposition~\ref{prop:classif}
if and only if all applications of Rule~R yield networks with $\lambda = 1$
relative to the original~\cite{leandro2009}.
When a neuron is not representable, the candidate with the highest $\lambda$
provides the best logical approximation.

\paragraph{Extraction procedure.}
Given a crystallized \LNN{}, symbolic extraction proceeds layer by layer,
bottom-up.
Each neuron is tested for direct representability (Proposition~\ref{prop:classif}).
If representable, the corresponding \L{}ukasiewicz formula is recorded.
If not, Rule~R is applied exhaustively and the best-$\lambda$ candidate
replaces the neuron.
The resulting formula is a \L{}ukasiewicz expression in the input variables.

% ── 2.4 Relationship to prior work ───────────────────────────────────────────
\subsection{Relationship to Leandro (2009)}
\label{ssec:prior}

The \LNN{} framework was established in~\cite{leandro2009}.
Compared to that work, the present paper introduces the following
novelties: (i)~a theoretical guarantee for residual blocks
(Theorem~\ref{thm:merge}) in place of the neuron classification of
Proposition~\ref{prop:classif}; (ii)~a corrected LM training procedure
(Section~\ref{ssec:lm}) that avoids the numerical instability present in
Leandro's original implementation; (iii)~two additional crystallization
strategies---STE and Proximal---applied to both flat and residual
architectures, whereas Leandro (2009) only experimented on MONK and small
UCI sets; and (iv)~a rigorous 5-way statistical comparison across six
datasets with ten trials each.

Proposition~\ref{prop:classif} and the rewriting algebra in
Sections~\ref{ssec:lnn}--\ref{ssec:rewriting} are reproduced from
\cite{leandro2009} as background; all material in
Sections~\ref{sec:residual}--\ref{sec:experiments} is new.

% ── 3. Residual Łukasiewicz Networks ─────────────────────────────────────────
\section{Residual \L{}ukasiewicz Networks}
\label{sec:residual}

The original \LNN{} framework is restricted to shallow architectures; residual
connections resolve this by introducing a structural constraint on the merge
neuron that guarantees Proposition~\ref{prop:classif} without any additional
training restriction.

\subsection{The \L{}ukasiewicz residual block}
\label{ssec:res_block}

\begin{definition}[\L{}ukasiewicz residual block]
\label{def:resblock}
Let $F^{(k)}\colon [0,1]^d \to [0,1]^d$ be a feed-forward linear layer
with real-valued weights during training at depth $k$.
The \emph{\L{}ukasiewicz residual block} at depth $k$ produces
\begin{equation}
  \mathbf{h}^{(k)} = \psi_{\mathbf{b}}\!\Bigl(
    F^{(k)}\!\bigl(\mathbf{h}^{(k-1)}\bigr)
    + \mathbf{h}^{(k-1)}
  \Bigr),
  \label{eq:resblock}
\end{equation}
where $\mathbf{b} \in \mathbb{R}^d$ is a vector of \emph{merge biases} that
are trained and crystallized alongside the weights of $F^{(k)}$, and
$\psi_{\mathbf{b}}$ denotes element-wise application of the scalar function
$\psi_{b_j}$ (Eq.~\eqref{eq:psib}) to the $j$-th coordinate.
\end{definition}

The output $\mathbf{h}^{(k)}_j$ is the result of a merge neuron that receives
two inputs: $F^{(k)}(\mathbf{h}^{(k-1)})_j$ and $\mathbf{h}^{(k-1)}_j$.
Both inputs arrive with weight $+1$ --- the residual path contributes
$F^{(k)}(\mathbf{h}^{(k-1)})_j$ with coefficient $+1$ and the shortcut
contributes $\mathbf{h}^{(k-1)}_j$ with coefficient $+1$.
The bias $b_j$ is the only free parameter of the merge neuron.

\subsection{Symbolic guarantee: Proposition~\ref{prop:classif} by construction}
\label{ssec:res_theorem}

The structural observation above leads directly to the main theoretical result
of this paper.

\begin{theorem}[Merge neurons satisfy Proposition~\ref{prop:classif} by construction]
\label{thm:merge}
Let $b_j \in \mathbb{Z}$ be the (crystallized) merge bias for component $j$
in a \L{}ukasiewicz residual block (Definition~\ref{def:resblock}).
Then the merge neuron for component $j$ satisfies Proposition~\ref{prop:classif}:
\begin{enumerate}
  \item If $b_j = -1$, the merge neuron computes the conjunction
        $F^{(k)}_j \tnorm h^{(k-1)}_j$.
  \item If $b_j = 0$, the merge neuron computes the disjunction
        $F^{(k)}_j \disj h^{(k-1)}_j$.
\end{enumerate}
No constraint on the weights of $F^{(k)}$ is required.
\end{theorem}

\begin{proof}
The merge neuron for component $j$ computes
$\psi\bigl(F^{(k)}_j + h^{(k-1)}_j + b_j\bigr)$
with input weights $(+1, +1)$.
Setting $n = 0$ (zero negative weights) and $p = 2$ (two positive weights)
in Proposition~\ref{prop:classif}:
\begin{itemize}
  \item Case 1 ($b_j = -p + 1 = -1$):
    $\psi(F^{(k)}_j + h^{(k-1)}_j - 1)
     = \max(0, F^{(k)}_j + h^{(k-1)}_j - 1)
     = F^{(k)}_j \tnorm h^{(k-1)}_j$.\quad$\checkmark$
  \item Case 2 ($b_j = n = 0$):
    $\psi(F^{(k)}_j + h^{(k-1)}_j)
     = \min(1, F^{(k)}_j + h^{(k-1)}_j)
     = F^{(k)}_j \disj h^{(k-1)}_j$.\quad$\checkmark$
\end{itemize}
Both cases follow from the definitions in Eqs.~\eqref{eq:tnorm}
and~\eqref{eq:disj}.
The merge bias values $b_j \in \{-1, 0\}$ are the only integer values
consistent with $n=0, p=2$ in Proposition~\ref{prop:classif};
all other integer biases would produce constant outputs on $[0,1]^2$
(e.g.\ $b_j = -2$: $\psi(x+y-2) = 0$ for all $x,y \in [0,1]$;
$b_j = 1$: $\psi(x+y+1) = 1$ for all $x,y \in [0,1]$).
\end{proof}

\begin{remark}
The proof is a direct application of Proposition~\ref{prop:classif} to the
specific parameterisation $n=0$, $p=2$ imposed by the residual block
structure.
The contribution of Theorem~\ref{thm:merge} is the architectural observation
that incorporating ResNet skip connections into \LNN{}s forces exactly this
parameterisation at the merge neuron, establishing the symbolic guarantee
without any additional training constraint.
\end{remark}


\begin{corollary}
In a \L{}ukasiewicz residual network, every merge neuron is
interpretable as either conjunction or disjunction, independently of
how well the inner layer $F^{(k)}$ has crystallized.
\end{corollary}

This result has a practical consequence: even when the inner layers of the
residual block contain neurons that are not directly representable by
Proposition~\ref{prop:classif} (requiring the rewriting algebra for
approximation), the merge neurons are guaranteed to be exact.
The interpretability of the network's \emph{skeleton} (the residual connections
and merge operations) is preserved by construction.

\subsection{Formula extraction from residual networks}
\label{ssec:res_extract}

Extraction from a residual \LNN{} proceeds by the same bottom-up procedure
described in Section~\ref{ssec:rewriting}, with the following addition.
At each residual block, the merge neuron is extracted first using
Theorem~\ref{thm:merge}: if the crystallized bias $b_j \in \{-1,0\}$, the
merge formula is $F^{(k)}_j \tnorm h^{(k-1)}_j$ or
$F^{(k)}_j \disj h^{(k-1)}_j$, respectively.
The sub-formula $F^{(k)}_j$ is then extracted from the inner layer neurons
using Proposition~\ref{prop:classif} (with the rewriting algebra for
non-representable cases).

As a concrete illustration, the residual \LNN{} trained on the MONK-3 problem
(Section~\ref{sec:experiments}) converges to the exact formula
\begin{equation}
  F = \neg\!\bigl(\mathit{body\_shape} \disj \mathit{jacket\_colour}\bigr),
  \label{eq:monk3}
\end{equation}
with the merge neuron taking $b_j = -1$ (conjunction of the shortcut and
the hidden-layer output) and the inner layer computing
$\mathit{body\_shape} \disj \mathit{jacket\_colour}$.
The formula achieves $97.2\%$ classification accuracy, correctly capturing the
ground truth rule.

% ── 4. Crystallization Strategies ────────────────────────────────────────────
\section{Crystallization Strategies}
\label{sec:crystal}

Proposition~\ref{prop:classif} requires all weights to be elements of
$\{-1, 0, +1\}$.
Gradient-based training evolves weights continuously in $\mathbb{R}$; the
process of driving them to integer values is called \emph{crystallization}.
We measure crystallization quality by the \emph{representation error}
\begin{equation}
  \Delta(N) = \sum_{k} \bigl(w_k - \lfloor w_k \rceil\bigr)^2,
  \label{eq:delta}
\end{equation}
where $\lfloor \cdot \rceil$ denotes rounding to the nearest integer.
A network is crystallized when $\Delta(N) < \varepsilon$ for a small
threshold $\varepsilon$ (we use $\varepsilon = 10^{-3}$ throughout).

\subsection{Smooth crystallization}
\label{ssec:smooth}

Hard rounding destroys differentiability and prevents gradient flow.
The \emph{smooth crystallization function}
\begin{equation}
  \Upsilon_n(w) = \operatorname{sgn}(w)\!\left(
    \cos^n\!\!\left((1 - \{|w|\})\tfrac{\pi}{2}\right) + \lfloor|w|\rfloor
  \right),
  \label{eq:upsilon}
\end{equation}
where $\{|w|\} = |w| - \lfloor|w|\rfloor$ is the fractional part,
deforms the identity toward the nearest integer while remaining
differentiable everywhere.
The exponent $n$ controls the strength of attraction: for $n = 2$, the
function is a gentle cosine deformation; as $n \to \infty$, it converges
to hard rounding.

A \emph{progressive crystallization schedule} applies $\Upsilon_n$ with
increasing $n$ over the final training epochs:
$\Upsilon_2 \to \Upsilon_4 \to \Upsilon_8 \to \Upsilon_{16}$.
This schedule avoids premature commitment (early rounding can destroy
gradient information) while ensuring that all quasi-integer weights are
driven to the correct integer before the final round.
At termination, weights are clamped to $[-1, +1]$ and rounded; biases
are rounded but not clamped, preserving multi-input conjunctions and
disjunctions.

\paragraph{Interaction with STE.}
For the STE strategy (Section~\ref{ssec:ste}), the smooth schedule is
applied to the continuous shadow weights $w_c$ between update steps during
the final training epochs: each $w_c$ is replaced by $\Upsilon_n(w_c)$
before the hard quantization $T(\Upsilon_n(w_c))$ is computed for the next
forward pass.
This ensures that shadow weights enter the final rounding in a state
that is already close to the target integer, reducing the gradient variance
introduced by the STE approximation near the quantization boundaries.

\subsection{Levenberg--Marquardt training}
\label{ssec:lm}

The Levenberg--Marquardt algorithm~\cite{levenberg1944,marquardt1963} is a
second-order method that interpolates between gradient descent and the
Gauss--Newton method.
Given residuals $\mathbf{e} = f_\theta(X) - y$ and Jacobian
$J = \partial \mathbf{e}/\partial w$, the weight update is
\begin{equation}
  \Delta w = -\bigl[J^\top J + \mu I\bigr]^{-1} J^\top \mathbf{e},
  \label{eq:lm}
\end{equation}
where $\mu > 0$ is the Levenberg damping parameter.

\paragraph{A critical correction.}
The original implementation in~\cite{leandro2009} used the Marquardt scaling
$\mu\,\operatorname{diag}(J^\top J)$ in place of $\mu I$.
When weights are near zero at initialization, the diagonal entries of
$J^\top J$ are also near zero; the system becomes ill-conditioned.
To quantify: if $\|w\| \approx \epsilon \ll 1$, the Jacobian entries
$J_{ij} = \partial e_i/\partial w_j$ are $O(\epsilon)$ because the
truncated-identity activation is nearly linear near zero, giving
$\mathrm{diag}(J^\top J) \sim O(\epsilon^2)$.
The Marquardt-damped system has effective regularization
$(1+\mu)\,\mathrm{diag}(J^\top J) \approx \mu\epsilon^2 I$, yielding
$\|\Delta w\| \approx \|\mathbf{e}\|/(\mu\epsilon^2)$.
For $\mu = 0.01$, $\epsilon = 0.1$, $\|\mathbf{e}\| = O(1)$, this gives
$\|\Delta w\| \sim 100$, consistent with the divergent magnitudes
($\|\Delta w\| \approx 47$) observed in practice at the typical
initialization scale.
The Levenberg form in Eq.~\eqref{eq:lm} uses the identity matrix, guaranteeing
$\|\Delta w\| \leq \|\mathbf{e}\|/\mu$ regardless of the weight magnitude.
All experiments in this paper use the corrected update.

\paragraph{Properties.}
LM is a second-order method and converges rapidly near a solution:
on structured small-scale problems (the four MONK datasets), the corrected
LM achieves crystallization in a mean of 6.2 iterations on MONK-3.
The crystallization guarantee is, however, \emph{probabilistic}: the
algorithm converges to a continuous optimum, which may or may not lie
sufficiently close to the integer lattice within the allocated iteration
budget.
With residual connections, the structural constraint on merge neurons
(Theorem~\ref{thm:merge}) reduces the number of weights that must
independently crystallize, improving the empirical crystallization rate
substantially (Section~\ref{sec:experiments}).

\subsection{Straight-Through Estimation}
\label{ssec:ste}

The straight-through estimator (STE), introduced by Bengio
et~al.~\cite{bengio2013}, resolves the differentiability problem of
weight quantization through a deliberate gradient approximation.

During the forward pass, each continuous weight $w_c \in \mathbb{R}$ is
quantized to the nearest element of $\{-1, 0, +1\}$ by
\begin{equation}
  T(w_c) = \operatorname{sgn}(w_c) \cdot \mathbf{1}\!\left[|w_c| > \tfrac{1}{3}\right].
  \label{eq:T}
\end{equation}
The threshold $\tfrac{1}{3}$ implements the minimum-distortion uniform
ternary partition of $[-1,+1]$: it divides the interval into three
equal-width regions
$(-\infty,-\tfrac{1}{3}) \to -1$,
$[-\tfrac{1}{3},+\tfrac{1}{3}] \to 0$,
$(\tfrac{1}{3},+\infty) \to +1$,
which is the canonical symmetric codebook for ternary quantization with a
uniform prior on $[-1,+1]$.
The network computes with $T(w_c)$ but stores and updates the continuous
shadow weight $w_c$.
During the backward pass, the gradient of the loss with respect to $w_c$ is
approximated by passing it directly through $T$ as if $T$ were the identity:
\begin{equation}
  \frac{\partial \mathcal{L}}{\partial w_c}
    \approx \frac{\partial \mathcal{L}}{\partial T(w_c)}.
  \label{eq:ste}
\end{equation}
The approximation is biased but has been shown to work in practice for
binarized and ternarized networks~\cite{courbariaux2016,li2016}: the
continuous weights learn to position themselves in regions where $T$ is
locally constant, and the direction of the approximate gradient is
sufficiently informative.

\paragraph{Crystallization guarantee.}
By construction, the final integer weights are $T(w_c)$ for each parameter.
Crystallization is $100\%$ guaranteed regardless of the problem structure
or dimensionality.
The cost is that all $n$ parameters are used (no sparsity), producing
networks whose inner neurons may not directly satisfy
Proposition~\ref{prop:classif}; the rewriting algebra (Section~\ref{ssec:rewriting})
is then required to extract a logical formula.

Applied to \texttt{LukResidualNet}, STE requires no algorithmic modification:
inner layers carry \texttt{mode='ste'} parameters quantized by $T(\cdot)$;
merge biases are treated as continuous shadow weights and rounded after training.
By Theorem~\ref{thm:merge}, crystallized merge biases $b_j \in \{-1,0\}$
yield exact conjunction or disjunction regardless of inner-layer representability,
providing a learnable pass-through that benefits datasets whose ground-truth
logic is expressible by a shallow formula.

\subsection{Proximal regularization}
\label{ssec:proximal}

The proximal strategy trains with Adam~\cite{kingma2015} on a composite
objective that jointly minimizes prediction error and drives weights toward
integer values:
\begin{equation}
  \mathcal{L} = \mathrm{MSE}
    + \lambda_s \sum_k |w_k|
    + \lambda_a \sum_k w_k^2(1 - w_k^2).
  \label{eq:proximal}
\end{equation}
The first penalty, $\lambda_s \sum|w_k|$, is an $\ell_1$ regularizer that
encourages sparsity by driving weights toward zero.
The second penalty, $\lambda_a \sum w_k^2(1-w_k^2)$, is the \emph{ternary
attraction potential}~\cite{parikh2014}: the function $P(w) = w^2(1-w^2)$
vanishes exactly at $w \in \{-1, 0, +1\}$ and achieves its maximum at
$|w| = 1/\sqrt{2} \approx 0.71$, creating a potential well around each
integer target.

Training proceeds in two phases.
In \emph{Phase~1} ($65\%$ of the iteration budget), only the MSE term is
active, allowing the network to find a continuous optimum without
interference from the regularizers.
The $65/35$ split was selected so that Phase~1 is long enough for the
network to locate a good loss basin before crystallization pressure is
applied; starting the regularizers too early (e.g.\ $50/50$) causes the
$\ell_1$ term to collapse weights before the network has converged, while
too late a start (e.g.\ $80/20$) leaves insufficient budget for reliable
crystallization.
In \emph{Phase~2} (the remaining $35\%$), $\lambda_s$ and $\lambda_a$ are
introduced gradually and then amplified by a factor of $10\times$ in the
final segment, pushing all near-integer weights to their nearest integer.
Crystallization is guaranteed by construction, as with STE.

\paragraph{Sparsity and failure mode.}
The $\ell_1$ penalty has the beneficial effect of driving many weights exactly
to zero, producing sparse networks: the Heart Disease experiment (Section~\ref{ssec:heart})
extracts a formula using only 5 of the available 22 input features.
The same penalty creates a failure mode on high-dimensional dense problems:
the $\ell_1$ regularizer eliminates discriminative weights before
the ternary attraction potential can stabilize a non-trivial solution,
collapsing the network to a constant-output classifier.
This failure manifests as $F_1 \approx 0$ and $\pm 0$ standard deviation across
all restarts (Section~\ref{sec:experiments}).

Applied to \texttt{LukResidualNet}, the proximal objective and two-phase
schedule are unchanged; Theorem~\ref{thm:merge} guarantees merge-bias
crystallization independently.
The residual shortcut does not mitigate the $\ell_1$ collapse:
Proximal-Res fails on the same five of six datasets as flat Proximal
(Section~\ref{sec:experiments}), confirming that the collapse is driven by
the $\ell_1$-gradient interaction rather than by signal loss through depth.

\subsection{Theoretical comparison}
\label{ssec:comparison}

Table~\ref{tab:theory} summarizes the theoretical properties of the three strategies.

\begin{table}[h]
\centering
\caption{Theoretical properties of the three crystallization strategies.
``Crystallization guaranteed'': all trials achieve $\Delta(N) < \varepsilon$.
``Sparsity'': tendency to produce zero weights.
``Exact Prop.~\ref{prop:classif}'': extracted formula satisfies the proposition
exactly without rewriting-algebra approximation.}
\label{tab:theory}
\setlength{\tabcolsep}{5pt}
\begin{tabular}{lcccc}
\toprule
Strategy & Arch. & Cryst.\ guar. & Sparsity & Merge exact \\
\midrule
LM    & Flat     & Probabilistic & No  & N/A \\
LM    & Residual & Probabilistic & No  & Yes  \\
STE              & Flat     & Yes           & No  & N/A \\
STE              & Residual & Yes           & No  & Yes \\
Proximal         & Flat     & Yes           & Yes & N/A \\
Proximal         & Residual & Yes           & Yes & Yes \\
\bottomrule
\end{tabular}
\end{table}

Every residual configuration benefits from Theorem~\ref{thm:merge}:
merge neurons satisfy Proposition~\ref{prop:classif} by construction,
regardless of the inner-layer training dynamics.
For the residual variants, the ``Merge exact'' column captures this
architectural guarantee; the inner neurons may still require the rewriting
algebra for complete formula extraction.
LM with flat architecture omits the ``Merge exact'' column because it
produces no merge neurons.

\subsection{Training and extraction pipeline}
\label{ssec:pipeline}

Algorithm~\ref{alg:pipeline} summarises the complete workflow for any of the
three strategies.

\begin{algorithm}
\caption{\LNN{} training and symbolic extraction}
\label{alg:pipeline}
\begin{algorithmic}[1]
\REQUIRE Dataset $(X, y)$; strategy $s \in \{\mathrm{LM\text{-}Res},
         \mathrm{STE}, \mathrm{STE\text{-}Res},
         \mathrm{Proximal}, \mathrm{Prx\text{-}Res}\}$; budget $T$;
         threshold $\varepsilon$
\ENSURE \L{}ukasiewicz formula $F$ or best-$\lambda$ approximation
\STATE Initialise $w \leftarrow$ small random values; $b \leftarrow 0$
\FOR{$t = 1$ \TO $T$}
  \STATE \textbf{Train} with strategy $s$:
         LM-Res via Eq.~\eqref{eq:lm};
         STE/STE-Res via Eqs.~\eqref{eq:T},\eqref{eq:ste};
         Proximal/Prx-Res via Eq.~\eqref{eq:proximal} with two-phase schedule
  \IF{$t > 0.85\,T$}
    \STATE Apply smooth schedule $\Upsilon_2 \!\to\! \Upsilon_{16}$
           to shadow weights $w_c$
  \ENDIF
\ENDFOR
\STATE Clamp $w_c \leftarrow \mathrm{clip}(w_c, -1, +1)$; round all $w_c$
       and $b$ to nearest integer
\STATE Compute $\Delta(N)$ (Eq.~\eqref{eq:delta}); crystallized iff
       $\Delta(N) < \varepsilon$
\STATE \textbf{Symbolic extraction} (bottom-up):
\FOR{each neuron $\alpha$ in topological order}
  \IF{$\alpha$ is a merge neuron with $b_\alpha \in \{-1, 0\}$}
    \STATE Apply Theorem~\ref{thm:merge} (exact conjunction / disjunction)
  \ELSIF{$\alpha$ satisfies Proposition~\ref{prop:classif}}
    \STATE Record \L{}ukasiewicz formula for $\alpha$
  \ELSE
    \STATE Apply Rule~R exhaustively; keep candidate with highest $\lambda$
  \ENDIF
\ENDFOR
\RETURN Composed formula $F$ over input variables
\end{algorithmic}
\end{algorithm}

% ── 5. Experimental Evaluation ───────────────────────────────────────────────
\section{Experimental Evaluation}
\label{sec:experiments}

\subsection{Protocol}
\label{ssec:protocol}

\paragraph{Datasets.}
We evaluate on six UCI~\cite{uci2017} benchmarks representing a range of
problem sizes and structures (Table~\ref{tab:datasets}).

\begin{table}[h]
\centering
\caption{Dataset characteristics. Class balance is the fraction of the majority class
(test set). $^{\dagger}$Mushroom training data augmented with synthetic half-valued
negative instances (edible mushrooms rescaled to $0.5$); enriched balance shown.}
\label{tab:datasets}
\setlength{\tabcolsep}{5pt}
\begin{tabular}{lrrrll}
\toprule
Dataset & Train & Test & Features & Class balance  \\
\midrule
Mushroom$^{\dagger}$ & 9\,866 & 2\,466 & 111 & $65.9\%$ \\
Heart Disease        & 243    & 60     & 22  & $54.1\%$ \\
MONK-1               & 124    & 432    & 17  & $50.0\%$ \\
MONK-2               & 169    & 432    & 17  & $67.1\%$ \\
MONK-3               & 122    & 432    & 17  & $52.8\%$ \\
Breast Cancer        & 229    & 57     & 15  & $70.3\%$ \\
\bottomrule
\end{tabular}
\end{table}

The Heart Disease dataset (Cleveland subset) uses 13 raw attributes;
five continuous features are scaled to $[0,1]$ via min-max normalization,
three binary attributes are kept as-is, four categorical attributes are
one-hot encoded, and one ordinal attribute (\texttt{ca}) is normalized to
$[0,1]$, yielding 22 input features.
MONK datasets use nominal attributes encoded as one-hot binary vectors;
MONK-3 includes $5\%$ attribute noise~\cite{thrun1991}.

\paragraph{Training setup.}
For each dataset and configuration, we run 10 independent trials with different
random seeds and select hyperparameters (architecture, learning rate, and
regularization strengths) by grid search on a validation split, reporting
test set results.
The five configurations are:
LM with residual connections (LM-Res), STE with flat architecture (STE),
STE with residual connections (STE-Res), Proximal with flat architecture
(Proximal), and Proximal with residual connections (Prx-Res).
LM is evaluated only with residual architecture; the flat LM is the original
method of~\cite{leandro2009} (its corrected form was already evaluated in the
prior conference version of this work).
All experiments use the smooth crystallization schedule of
Section~\ref{ssec:smooth}.
Statistical comparisons use the Wilcoxon signed-rank test~($n = 10$, $\alpha = 0.05$).
On synthetic truth tables, LM achieves exact recovery when crystallization
succeeds ($F_1 = 0.720$ on a two-clause formula, 5 of 10 trials); Proximal
collapses to constant output on the same formula (all 10 trials), confirming
the failure mode of Section~\ref{ssec:proximal}.

\paragraph{Evaluation metrics.}
We report mean accuracy and $F_1$ score over 10 trials, with standard
deviations, together with crystallization rate (fraction of trials achieving
$\Delta(N) < 10^{-3}$).
Training time (wall clock, seconds) is reported separately.

\subsection{Crystallization reliability on real datasets}
\label{ssec:cryst_real}

Proximal and Prx-Res are included as structural negative controls: their
collapse is reproducible across all hyperparameter configurations and
theoretically explained in Section~\ref{ssec:proximal}.
Tables~\ref{tab:acc} and~\ref{tab:f1} report accuracy, $F_1$, and
crystallization rates for all five configurations across six datasets.
STE, STE-Res, Proximal, and Prx-Res achieve $10/10$ crystallization on every
dataset by construction.
LM-Res crystallizes reliably only when a near-integer optimum is reachable
within the iteration budget: $10/10$ on MONK-3, $9/10$ on Heart Disease,
$7/10$ on MONK-2, $6/10$ on Mushroom, $5/10$ on MONK-1, and $0/10$ on Breast
Cancer.

The Heart Disease result ($9/10$ for LM-Res) is noteworthy: the flat LM
without residual connections achieves $0/10$ crystallization on the same
problem~\cite{leandro2009}.
The merge-neuron constraint of Theorem~\ref{thm:merge} reduces the effective
number of weights that must independently crystallize, making the problem more
tractable for the second-order solver.
This improvement is architecture-specific to LM; the guaranteed-crystallization
methods (STE, Proximal, and their residual variants) do not benefit from this
reduction since crystallization is enforced by design in all trials.

\subsection{Classification accuracy and \texorpdfstring{$F_1$}{F1}}
\label{ssec:accuracy}

\begin{table*}[t]
\centering
\caption{Mean accuracy ($\pm$ std, 10 trials) and crystallization rate for all
five configurations.
$^{\ast}$: Wilcoxon $p < 0.05$ (STE-Res\,$>$\,Prx-Res).
Best accuracy per dataset in \textbf{bold}.
$^{\dagger}$: LM-Res achieved $0/10$ crystallization on Breast Cancer;
accuracy is from the uncrystallized continuous model (no formula extracted).}
\label{tab:acc}
\setlength{\tabcolsep}{4pt}
\begin{tabular}{l ccccc @{\hspace{8pt}} ccccc}
\toprule
 & \multicolumn{5}{c}{Accuracy (mean $\pm$ std)} &
   \multicolumn{5}{c}{Crystallization rate} \\
\cmidrule(lr){2-6}\cmidrule(lr){7-11}
Dataset & LM-Res & STE & STE-Res & Proximal & Prx-Res
        & LM-Res & STE & STE-Res & Proximal & Prx-Res \\
\midrule
Mushroom
  & $\mathbf{.668\pm.028}$ & $.665\pm.040$ & $.626\pm.102$
  & $.660\pm.000$          & $.660\pm.000$
  & $6/10$ & $10/10$ & $10/10$ & $10/10$ & $10/10$ \\
Heart Disease
  & $.592\pm.122$ & $\mathbf{.607\pm.101}$ & $.558\pm.088$
  & $.600\pm.000$ & $.600\pm.000$
  & $9/10$ & $10/10$ & $10/10$ & $10/10$ & $10/10$ \\
MONK-1$^{\ast}$
  & $.515\pm.296$ & $.600\pm.125$ & $\mathbf{.674\pm.118}$
  & $.558\pm.125$ & $.500\pm.000$
  & $5/10$ & $10/10$ & $10/10$ & $10/10$ & $10/10$ \\
MONK-2
  & $.631\pm.243$ & $\mathbf{.704\pm.070}$ & $.656\pm.145$
  & $.671\pm.000$ & $.671\pm.000$
  & $7/10$ & $10/10$ & $10/10$ & $10/10$ & $10/10$ \\
MONK-3$^{\ast}$
  & $.714\pm.212$ & $.681\pm.158$ & $\mathbf{.717\pm.149}$
  & $.472\pm.000$ & $.472\pm.000$
  & $10/10$ & $10/10$ & $10/10$ & $10/10$ & $10/10$ \\
Breast Cancer
  & $\mathbf{.632\pm.000}^{\dagger}$ & $.572\pm.120$ & $.605\pm.087$
  & $\mathbf{.632\pm.000}$ & $\mathbf{.632\pm.000}$
  & $0/10$ & $10/10$ & $10/10$ & $10/10$ & $10/10$ \\
\bottomrule
\end{tabular}
\smallskip\\
{\footnotesize $^{\ast}$MONK-1: $p(\text{STE-Res} > \text{Prx-Res}) = 0.008$;
MONK-3: $p(\text{STE-Res} > \text{Prx-Res}) = 0.004$ (Wilcoxon, $n = 10$).}
\end{table*}

\begin{table*}[t]
\centering
\caption{Mean $F_1$ score ($\pm$ std, 10 trials).
$^{\ast}$: Wilcoxon $p < 0.05$.
Best $F_1$ per dataset in \textbf{bold}.}
\label{tab:f1}
\setlength{\tabcolsep}{5pt}
\begin{tabular}{l ccccc}
\toprule
Dataset & LM-Res & STE & STE-Res & Proximal & Prx-Res \\
\midrule
Mushroom
  & $.103\pm.223$ & $.229\pm.316$ & $\mathbf{.299\pm.298}$
  & $.000\pm.000$ & $.000\pm.000$ \\
Heart Disease
  & $.462\pm.325$ & $\mathbf{.482\pm.337}$ & $.272\pm.330$
  & $.000\pm.000$ & $.000\pm.000$ \\
MONK-1$^{\ast}$
  & $.513\pm.263$ & $.481\pm.337$ & $\mathbf{.648\pm.240}$
  & $.147\pm.311$ & $.000\pm.000$ \\
MONK-2$^{\ast}$
  & $.366\pm.347$ & $.179\pm.286$ & $\mathbf{.455\pm.267}$
  & $.000\pm.000$ & $.000\pm.000$ \\
MONK-3$^{\ast}$
  & $.603\pm.425$ & $.612\pm.331$ & $\mathbf{.716\pm.259}$
  & $.000\pm.000$ & $.000\pm.000$ \\
Breast Cancer
  & $.000\pm.000$ & $.109\pm.230$ & $\mathbf{.137\pm.196}$
  & $.000\pm.000$ & $.000\pm.000$ \\
\bottomrule
\end{tabular}
\smallskip\\
{\footnotesize $^{\ast}$MONK-1: $p(\text{STE-Res} > \text{Prx-Res})_{F_1} = 0.004$;
MONK-2: $p(\text{STE} < \text{STE-Res})_{F_1} = 0.020$,
$p(\text{STE-Res} > \text{Prx-Res})_{F_1} = 0.004$;
MONK-3: $p(\text{STE-Res} > \text{Prx-Res})_{F_1} = 0.004$;
Mushroom: $p(\text{STE-Res} > \text{Prx-Res})_{F_1} = 0.016$.}
\end{table*}

No single configuration dominates across all datasets.
STE-Res achieves the highest accuracy on MONK-1 ($0.674$, $+12.4$~pp over
flat STE) and MONK-3 ($0.717$, $+3.6$~pp over STE), narrowly edging out
LM-Res ($0.714$) on MONK-3.
STE leads on MONK-2 ($+4.8$~pp over STE-Res) and Heart Disease.
LM-Res leads on Mushroom ($0.668$).

The $F_1$ picture is clearer: STE-Res achieves the highest mean $F_1$ on four
of six datasets (MONK-1: $0.648$, MONK-2: $0.455$, MONK-3: $0.716$, Mushroom:
$0.299$), with the differences on MONK-2 and MONK-3 statistically significant
(Wilcoxon $p = 0.020$ and $p = 0.004$ respectively vs flat STE; all four
STE-Res vs Prx-Res gaps are significant at $p \leq 0.016$).
The $F_1$ improvement on Heart Disease is an exception: STE-Res ($0.272$) lags
behind STE ($0.482$), suggesting that the residual connections hinder
performance on this small, low-dimensional problem.
The accuracy differences between STE and STE-Res are not statistically
significant on any individual dataset (all accuracy Wilcoxon $p > 0.05$), but
the consistent direction of the $F_1$ advantage on three of four structured
MONK datasets supports the hypothesis that residual connections benefit
problems with underlying logical structure.

The degenerate Proximal solutions on five of six datasets (accuracy equals
majority-class rate with $\pm 0.000$ standard deviation, $F_1 = 0$) confirm
the failure mode as structural: grid search was conducted over a wide range of
$\lambda_s$ and $\lambda_a$ values, and collapse persists regardless of
architecture.
Prx-Res shows identical degenerate behaviour to flat Proximal on all five
affected datasets, confirming that the residual shortcut does not mitigate the
$\ell_1$ collapse.
The single non-degenerate Proximal result is flat Proximal on MONK-1 ($F_1 =
0.147$); Prx-Res on MONK-1 collapses completely.

\subsection{Formula quality and interpretability}
\label{ssec:formulas}

Across all experiments, the only fully representable formula in the sense of
Proposition~\ref{prop:classif} --- a formula extractable without any
rewriting-algebra approximation --- emerges from LM-Res on MONK-3.

\paragraph{MONK-3 exact formula.}
In trial~1, LM-Res converges in $6$ iterations to the fully crystallized
network with accuracy $97.2\%$.
The extracted formula is:
\begin{equation}
  F = \neg\!\bigl(\mathit{body\_shape} \disj \mathit{jacket\_colour}\bigr),
  \label{eq:monk3formula}
\end{equation}
which classifies all but $2.8\%$ of the test set (the erroneous labels
introduced by $5\%$ noise).
All neurons satisfy Proposition~\ref{prop:classif} exactly: the inner layer
computes $\mathit{body\_shape} \disj \mathit{jacket\_colour}$ ($b=0$, case~2),
the merge neuron computes $y \tnorm y$ ($b_j=-1$, case~1), and on binary inputs
$y \tnorm y = y$, so the output negation closes the formula.
STE and Proximal also crystallize on MONK-3 (10/10) but require the rewriting
algebra for inner neurons.

\paragraph{Heart Disease formula extraction.}
\label{ssec:heart}
We highlight a Proximal trial that illustrates the framework's capability for
clinically interpretable rule extraction.
This trial was conducted with manually selected hyperparameters outside the
grid search; the systematic grid search (Section~\ref{ssec:accuracy}) finds
that Proximal collapses to a majority-class solution on Heart Disease across
all tested hyperparameter combinations.
The trial below therefore illustrates a \emph{best-case capability} of the
Proximal strategy rather than its typical behaviour on this dataset.
After $321$ iterations, the trial produces a maximally sparse
crystallized network on the Heart Disease dataset.
Four of the six first-layer neurons carry only zero weights and are discarded
by the symbolic extractor.
The two active neurons are:
\begin{equation}\label{eq:n1}
\begin{split}
  n_1 &= \psi_{+1}(-\mathtt{trestbps},\; -\mathtt{oldpeak},\; -\mathtt{ca}) \\
      &= \neg\mathtt{trestbps} \tnorm \neg\mathtt{oldpeak} \tnorm \neg\mathtt{ca},
\end{split}
\end{equation}
\begin{equation}\label{eq:n5}
\begin{split}
  n_5 &= \psi_{0}(+\mathtt{thalach},\; -\mathtt{ca},\; -\mathtt{cp}{=}4) \\
      &= \mathtt{thalach} \tnorm \neg\mathtt{ca} \tnorm \neg(\mathtt{cp}{=}4).
\end{split}
\end{equation}
where the bias conditions $b = -p+1 = +1$ in $n_1$ and $b = n = 0$ in $n_5$
both satisfy Proposition~\ref{prop:classif} (conjunctions with three inputs).
The output layer combines $n_1$ and $n_5$ by:
$i_1 = \psi_{+1}(-n_1, -n_5) = \neg n_1 \tnorm \neg n_5$.
Substituting yields the complete formula:
\begin{equation}
  \boxed{
    \begin{aligned}
      \mathrm{DISEASE} =&\;
      \neg\!\bigl(\neg\mathtt{trestbps} \tnorm \neg\mathtt{oldpeak}
                  \tnorm \neg\mathtt{ca}\bigr) \\
      &\qquad \tnorm
      \neg\!\bigl(\mathtt{thalach} \tnorm \neg\mathtt{ca}
                  \tnorm \neg(\mathtt{cp}{=}4)\bigr)
    \end{aligned}
  }
  \label{eq:heart_formula}
\end{equation}
($\mathrm{DISEASE} = 1$ indicates the presence of coronary artery disease).

\paragraph{Numerical evaluation.}
Table~\ref{tab:heart_trace} evaluates Eq.~\eqref{eq:heart_formula} for two
representative test patients using the \L{}ukasiewicz conjunction
$a \tnorm b = \max(0, a + b - 1)$.
Patient~A has complete vessel occlusion (\texttt{ca}$=1.00$), which zeroes
both $n_1$ and $n_5$, driving the output to $1.00$ (disease).
Patient~B presents with no stenosis, normal blood pressure, and good
exercise capacity; the output is $0.00$ (healthy).
A threshold of $0.5$ separates the two correctly.
The formula can be verified step by step by any reader who knows the single
operator definition $a \tnorm b = \max(0, a + b - 1)$.

\begin{table}[h]
\centering
\caption{Evaluation trace for Eq.~\eqref{eq:heart_formula} on two test
patients (features normalised to $[0,1]$).}
\label{tab:heart_trace}
\setlength{\tabcolsep}{5pt}
\begin{tabular}{lrr}
\toprule
 & Patient A & Patient B \\
  & (disease) & (healthy) \\
\midrule
\texttt{ca} & $1.00$ & $0.00$ \\
\texttt{trestbps} & $0.64$ & $0.27$ \\
\texttt{oldpeak} & $0.42$ & $0.00$ \\
\texttt{thalach} & $0.29$ & $0.79$ \\
\texttt{cp=4} indicator & $1$ & $0$ \\
\midrule
$\neg\mathtt{ca}$ & $0.00$ & $1.00$ \\
$\neg\mathtt{trestbps}$ & $0.36$ & $0.73$ \\
$\neg\mathtt{oldpeak}$ & $0.58$ & $1.00$ \\
$\neg(\mathtt{cp{=}4})$ & $0.00$ & $1.00$ \\
\midrule
$n_1 = \neg\mathtt{t} \tnorm \neg\mathtt{o} \tnorm \neg\mathtt{ca}$
  & $0.00$
  & $0.73$ \\
$n_5 = \mathtt{tal} \tnorm \neg\mathtt{ca} \tnorm \neg(\mathtt{cp{=}4})$
  & $ 0.00$
  & $ 0.79$ \\
\midrule
$\neg n_1$ & $1.00$ & $0.27$ \\
$\neg n_5$ & $1.00$ & $0.21$ \\
$\mathrm{DISEASE} = \neg n_1 \tnorm \neg n_5$
  & $\mathbf{1.00}$
  & $\mathbf{0.00}$ \\
\bottomrule
\end{tabular}
\end{table}
{\footnotesize $\mathtt{t}$=\texttt{trestbps}, $\mathtt{o}$=\texttt{oldpeak},
$\mathtt{tal}$=\texttt{thalach}. Values are illustrative from the test set.}

Only 5 of the 22 input features received non-zero weights
(Table~\ref{tab:heart_features}).
Every selected feature is an established predictor of coronary artery disease
in the cardiology literature~\cite{detrano1989,rudin2019}.
Most striking is the selection consistency: \texttt{ca} (the number of major
vessels with $>50\%$ stenosis) is the only feature selected in every one of
the 10 crystallized Proximal trials, regardless of random initialization.
We note that \texttt{ca} is universally recognized as the dominant predictor
in the Cleveland dataset~\cite{detrano1989}; the result is therefore not a
clinical discovery but a \emph{validation} that the Proximal extractor
reliably reproduces established medical knowledge, confirming that the
$\ell_1$ regularizer does not spuriously suppress clinically established
predictors under initialization variance.

\begin{table}[h]
\centering
\caption{Features selected in the Heart Disease formula (Eq.~\ref{eq:heart_formula}),
with clinical interpretation. ``Established'': recognized predictor
of coronary artery disease~\cite{detrano1989}.}
\label{tab:heart_features}
\setlength{\tabcolsep}{5pt}
\begin{tabular}{cll}
\toprule
Feature & Clinical meaning  \\
\midrule
\texttt{ca}       & Number of major vessels with $>50\%$ stenosis  \\
\texttt{trestbps} & Resting systolic blood pressure (mmHg)         \\
\texttt{oldpeak}  & ST-segment depression at exercise              \\
\texttt{thalach}  & Maximum heart rate achieved (bpm)              \\
\texttt{cp=4}     & Asymptomatic chest pain presentation           \\
\bottomrule
\end{tabular}
\end{table}

Evaluated on the held-out test set (61 patients: 37 healthy, 24 with disease),
formula~\eqref{eq:heart_formula} achieves accuracy $85.0\%$, sensitivity
$87.5\%$, specificity $83.3\%$, and $F_1 = 0.824$, identifying 21 of 24
disease cases correctly.
The formula can be verified analytically by any clinician: it classifies a
patient as diseased when neither the triple of low-risk indicators (normal
resting blood pressure, no ST depression, no blocked vessels) nor the triple
of adequate cardiac reserve indicators (high maximum heart rate, no blocked
vessels, symptomatic chest pain) are simultaneously satisfied.

\subsection{Training efficiency}
\label{ssec:time}

\begin{table*}[t]
\centering
\caption{Mean wall-clock training time (seconds, 10 trials).
Fastest configuration per dataset in \textbf{bold}.
Mean iteration counts are given in parentheses.}
\label{tab:time}
\setlength{\tabcolsep}{6pt}
\begin{tabular}{l rrrrr}
\toprule
Dataset & LM-Res & STE & STE-Res & Proximal & Prx-Res \\
\midrule
Mushroom
  & $10.24\ (255)$ & $88.71\ (461)$ & $368.80\ (1114)$
  & $6.88\ (44)$   & $\mathbf{4.54\ (24)}$ \\
Heart Disease
  & $2.25\ (96)$  & $1.00\ (175)$  & $2.79\ (459)$
  & $0.23\ (39)$  & $\mathbf{0.16\ (29)}$ \\
MONK-1
  & $6.34\ (143)$ & $\mathbf{0.39\ (306)}$ & $0.53\ (377)$
  & $0.57\ (217)$ & $0.44\ (83)$ \\
MONK-2
  & $4.77\ (135)$ & $3.30\ (482)$ & $2.46\ (372)$
  & $4.37\ (568)$ & $\mathbf{1.85\ (108)}$ \\
MONK-3
  & $1.25\ (6)$   & $0.30\ (253)$ & $\mathbf{0.21\ (154)}$
  & $0.35\ (49)$  & $0.37\ (43)$ \\
Breast Cancer
  & $15.42\ (102)$ & $3.29\ (561)$ & $4.23\ (719)$
  & $\mathbf{0.90\ (154)}$ & $0.93\ (160)$ \\
\bottomrule
\end{tabular}
\end{table*}

Training times are dominated by architecture-strategy interactions.
LM-Res is the fastest on MONK-3 ($1.25$~s, mean $6.2$ iterations) owing to
second-order convergence near the integer lattice, but is the slowest on Breast
Cancer ($15.4$~s): the Jacobian inversion has $O(p^2)$ cost per iteration,
and with $102$ mean iterations on a $20$-feature dataset, the per-iteration
cost dominates.
On Mushroom ($111$ features), LM-Res takes $10.2$~s while Prx-Res takes
$4.5$~s, reflecting the quadratic-versus-linear cost advantage of first-order
methods on high-dimensional problems.

STE-Res is substantially slower than flat STE: $368.8$~s versus $88.7$~s on
Mushroom (the residual architecture requires $1114$ vs $461$ mean iterations),
and $2.79$~s versus $1.00$~s on Heart Disease ($459$ vs $175$ iterations).
STE-Res converges fastest of all on MONK-3 ($0.21$~s, $154$ iterations),
where the residual pass-through helps reach a good solution quickly.

Proximal and Prx-Res are the fastest configurations on most datasets, because
Phase~1 (pure MSE minimization) terminates aggressively and the two-phase
schedule limits total iterations.
This speed advantage comes at a high cost in solution quality:
Proximal and Prx-Res produce degenerate solutions on five of six datasets
(Section~\ref{ssec:accuracy}), so the fast convergence reflects early
collapse rather than efficient learning.

% ── 6. Related Work ───────────────────────────────────────────────────────────
\section{Related Work}
\label{sec:related}

\subsection{Rule extraction from neural networks}

Extracting symbolic rules from trained neural networks has been studied since
the late 1980s.
Gallant~\cite{gallant1988} proposed connectionist expert systems in which
threshold units encode conjunctions; the extracted rules are Boolean.
KBANN~\cite{towell1994} injects domain knowledge into a network and refines it
through training, then extracts updated rules via a search over activation patterns.
Towell and Shavlik~\cite{towell1993} systematized this approach with a
weight-based extraction heuristic.
All these methods extract rules \emph{post-hoc}: the network is trained as a
continuous model and then approximated by a symbolic rule.
The approximation introduces an irreducible fidelity gap.

The \LNN{} approach is qualitatively different: the symbolic formula is not an
approximation to the network but an exact reading of it.
When all neurons satisfy Proposition~\ref{prop:classif}, the network \emph{is}
the formula.
When the rewriting algebra is needed, the approximation is bounded and
measurable by the $\lambda$-similarity metric rather than being assessed
informally.

\subsection{Fuzzy and many-valued neural systems}

ANFIS~\cite{jang1993} and related fuzzy neural architectures learn
membership functions and combine them with T-norm rules, producing interpretable
models that resemble \LNN{}s in spirit.
The key difference is that ANFIS membership functions are learned as continuous
parameters without a crystallization step; no exact correspondence between
neurons and logical connectives is established or enforced.
Amato~et~al.~\cite{amato2002} proved that neurons with rational weights can
represent rational \L{}ukasiewicz functions, extending the scope of the
Castro--Trillas result, but did not provide a training algorithm for integer
weights.
The present work provides such an algorithm (in three variants) and the
corresponding theoretical guarantees.

A contemporaneous line of work applies the \L{}ukasiewicz T-norm in a
different architectural setting.
Nguy and Wasilewski~\cite{nguy2025} study Differentiable Logic Gate Networks
(DLNs) in which each computational unit learns a probability distribution over
the 16 Boolean operations via softmax; the \L{}ukasiewicz T-norm
$T_L(a,b)=\max\{0,a+b-1\}$ replaces the Menger T-norm as the fuzzy relaxation.
Post-training, each unit is snapped to its most probable gate, converting the
DLN into a Binary Logic Gate Network (LGN) --- a process analogous to
crystallization.
To address vanishing gradients in deeper DLNs, they employ a residual
initialization scheme that biases units toward pass-through gates, an
engineering heuristic that parallels our use of skip connections but is not
backed by a structural theorem.
Three distinctions separate the two approaches: (i)~DLNs select \emph{which}
operation to apply at each node, whereas \LNN{}s fix the Łukasiewicz
semantics and crystallize \emph{connection weights}; (ii)~DLN discretization
yields logic-gate circuits without an explicit logical formula, while \LNN{}
crystallization directly exposes a readable \L{}ukasiewicz formula;
(iii)~the residual guarantee of Theorem~\ref{thm:merge} is a formal
structural property, whereas DLN residual initialization is a training
heuristic motivated by gradient flow.

\subsection{Weight quantization}

The STE used in Section~\ref{ssec:ste} was developed in the context of
binarized neural networks~\cite{bengio2013,courbariaux2016}, where weights are
restricted to $\{-1, +1\}$ for hardware efficiency.
Ternary weight networks~\cite{li2016} extend this to $\{-1, 0, +1\}$,
matching the target domain of crystallization.
The objective of quantization in these works is computational efficiency,
not logical interpretability: the quantized weights are not interpreted as
logical connectives, and no formula extraction procedure is defined.
The proximal objective of Section~\ref{ssec:proximal} is superficially related
to $\ell_1$-regularized training, but the ternary attraction term $P(w)$
is specific to the discrete target set $\{-1, 0, +1\}$ and is designed to
ensure that crystallization produces logically meaningful weights.

\subsection{Neuro-symbolic AI}

The broader neuro-symbolic AI programme~\cite{garcez2019,garcez2023} seeks to
combine neural learning with symbolic reasoning.
Riegel et al.~\cite{riegel2020} introduced Logical Neural Networks (IBM LNNs),
where every neuron represents a weighted real-valued logical operator; unlike
\LNN{}s, IBM LNN weights remain real-valued, no crystallization is defined, and
no neuron-to-connective correspondence is formally established.
Serafini and d'Avila Garcez~\cite{serafini2016} proposed Logic Tensor Networks
(LTNs), which ground first-order formulae in neural networks via real-valued
\L{}ukasiewicz t-norm satisfaction; LTN weights remain continuous and no
discrete formula is extracted.
Evans and Grefenstette~\cite{evans2018} learn first-order logic clauses by
differentiating through a discrete rule space rather than a neural architecture.
Manhaeve et al.~\cite{manhaeve2018} combine neural networks with probabilistic
logic programming; their uncertainty model (probability over logical facts)
complements rather than competes with \L{}ukasiewicz degrees of truth.
Unlike all of these, an \LNN{} after crystallization \emph{is} the formula:
network weights constitute the symbolic representation directly, and
Theorem~\ref{thm:merge} provides a structural guarantee unavailable in any of
the above frameworks.

% ── 7. Discussion ─────────────────────────────────────────────────────────────
\section{Discussion}
\label{sec:discussion}

\subsection{Interpretability by construction versus post-hoc}

The distinction between interpretability by construction and post-hoc
explanations matters most precisely when it is hardest to verify.
A post-hoc explanation is a separate model trained to approximate the
decisions of the original model; its fidelity to the original is measurable
but never guaranteed to be $1.0$~\cite{guidotti2018}.
In the \LNN{} framework, when a network is fully crystallized and every neuron
satisfies Proposition~\ref{prop:classif}, the formula \emph{is} the model ---
there is no approximation step, and the verification question reduces to
algebraic manipulation of the formula rather than statistical estimation.
Theorem~\ref{thm:merge} extends this guarantee to the residual merge neurons
unconditionally.

This property is particularly significant in safety-critical applications.
A clinical decision rule that can be written as a \L{}ukasiewicz formula can
be audited by domain experts working from the formula alone, without reference
to the training data or the training procedure.
The step-by-step evaluation in Table~\ref{tab:heart_trace} illustrates this:
a reader who knows only $a \tnorm b = \max(0, a+b-1)$ can verify the
classification of any patient independently.

\subsection{Architecture \texorpdfstring{$\times$}{×} strategy interaction}

The five-way comparison reveals that the benefit of residual connections is
strategy-dependent.
For STE, residual connections consistently improve $F_1$ on structured datasets
(MONK-1, MONK-2, MONK-3), where the underlying rule is logically compact.
The improvement is plausibly explained by the learnable pass-through property:
STE-Res can represent shallow functions by setting inner weights to zero, while
flat STE forces every activation through all layers.
For Proximal, residual connections provide no benefit: the $\ell_1$ collapse
dominates on the same datasets regardless of architecture.
This asymmetry suggests that the value of residual connections in \LNN{}s
depends critically on whether the crystallization strategy can effectively
utilise the pass-through path --- a path that Proximal's regularization
pressure systematically destroys.

\subsection{Comparison with DLN\textsubscript{\textit{L}} (Nguy and Wasilewski, 2025)}
\label{ssec:vs_dln}

Nguy and Wasilewski~\cite{nguy2025} report accuracy on MONK-1/2/3 and Breast
Cancer using their \L{}ukasiewicz-based DLN variants (100 runs each).
Table~\ref{tab:vs_dln} confronts those figures with our five configurations
(10 trials each).

\begin{table}[ht]
\centering
\caption{Test accuracy (\%) on overlapping datasets.
DLN$_L$, DLN$_{L^*}$ from Nguy and Wasilewski~\cite{nguy2025} (100~runs);
our results from 10~independent trials ($\bar{x}\pm\sigma$).
Best result per row in \textbf{bold}.}
\label{tab:vs_dln}
\setlength{\tabcolsep}{4pt}
\begin{tabular}{lcccccc}
\toprule
Dataset & DLN$_L$ & DLN$_{L^*}$ & LM-Res & STE & STE-Res \\
\midrule
MONK-1       & \textbf{100.0} & \textbf{100.0} & 51.5 & 60.0 & 67.4 \\
MONK-2       & 56.9           & 61.6           & 63.1 & \textbf{70.4} & 65.6 \\
MONK-3       & \textbf{94.0} & 78.2           & 71.4 & 68.1 & 71.7 \\
Breast Cancer & 70.0          & \textbf{74.3}  & 63.2 & 57.2 & 60.5 \\
\bottomrule
\end{tabular}
\end{table}

DLNs lead on MONK-1/3 (100\%/94.0\% vs.\ our 67.4\%/71.7\%), where compact
DNF formulas over binary features favour gate-selection directly.
\LNN{}s lead on MONK-2 (STE 70.4\% vs.\ DLN$_L$ 56.9\%), where a parity-like
concept is difficult to express as a small gate circuit.
On Breast Cancer, DLN$_{L^*}$ (74.3\%) outperforms STE-Res (60.5\%),
reflecting DLNs' tolerance for continuous features that need not converge to
ternary weights.
The two frameworks occupy complementary niches: DLNs are more accurate when
the target is a Boolean formula over binary features; \LNN{}s provide a richer
symbolic output --- a readable \L{}ukasiewicz formula with a formal structural
guarantee --- when interpretability by construction is the primary goal.

\subsection{Limitations}

Several limitations of the current framework should be acknowledged.
First, the LM solver has quadratic complexity in the number of parameters
($O(p^2)$ for the Jacobian inversion), making it impractical for very large
networks; the Mushroom experiment ($111$ features) already required mini-batch
Jacobian approximation~\cite{battiti1992}.
Second, the Proximal failure mode --- collapse to constant-output solutions ---
is structural and is not fully resolved by hyperparameter tuning or by
incorporating residual connections.
Feature pre-selection before Proximal training (using, for example, an initial
STE run) is a practical mitigation, but it adds a training step.
Third, the expressive power of \LNN{}s is bounded by the class of McNaughton
functions~\cite{mcnaughton1951,hajek1998}; it is not known whether every
classifiable decision boundary can be approximated to arbitrary accuracy by a
\LNN{} of bounded depth, and this question is open.
Fourth, this paper operates exclusively in the propositional \L{}ukasiewicz
framework; whether the residual block construction extends to the first-order
case (where quantifiers introduce cross-input dependencies) is an open question.

% ── 8. Conclusion ─────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}

The \L{}ukasiewicz neural network framework, established in~\cite{leandro2009},
rests on an exact correspondence between neurons with integer weights and
logical connectives.
This paper has extended the framework along two axes.

On the theoretical side, we proved that residual connections can be incorporated
into \LNN{}s with the symbolic guarantee preserved \emph{by construction}:
merge neurons in a \L{}ukasiewicz residual block always satisfy
Proposition~\ref{prop:classif}, regardless of the training procedure applied to
the inner layers.
On the practical side, we analysed three crystallization strategies (corrected
LM, STE, and proximal regularization), each evaluated in both flat and residual
architectures, and characterised the conditions under which each succeeds or
fails.

The five-way experimental evaluation across six benchmarks yields three
principal findings.
First, \emph{STE with residual connections} provides the best aggregate
discriminative quality: it achieves the highest mean $F_1$ on four of six
datasets, with significant improvements over Proximal-Res on MONK-1, MONK-2,
and MONK-3.
The improvement over flat STE is consistent in direction (MONK-1, MONK-2,
MONK-3) but not individually statistically significant, pointing to a positive
effect of the learnable pass-through in residual blocks.
Second, \emph{LM with residual connections} remains the most accurate method
on Mushroom and is the fastest on structured problems (6.2 mean iterations on
MONK-3), recovering exact logical formulas when crystallization succeeds.
Third, \emph{proximal regularization collapses} on five of six datasets
regardless of whether residual connections are used, confirming that the failure
mode is intrinsic to the $\ell_1$-dominated objective rather than to the
architecture.

The Heart Disease formula~\eqref{eq:heart_formula}, with five clinically
validated variables and a step-by-step verifiable structure, demonstrates that
the framework can produce transparent diagnostic rules from real medical data.
Together, these results establish \L{}ukasiewicz neural networks as a mature
framework for interpretable neuro-symbolic learning, with a principled
catalogue of integer-weight training strategies and a rigorous extension to
deeper, residual architectures.
Open directions include: whether a universal approximation theorem holds for
McNaughton functions under bounded-depth \LNN{}s; whether residuated-lattice
logics beyond \L{}ukasiewicz admit analogous neural encodings; and how
residual-block count affects inner-layer crystallization dynamics.

% ── Mandatory statements ──────────────────────────────────────────────────────
\section*{Data Availability Statement}

All datasets used in this study are publicly available from the UCI Machine
Learning Repository~\cite{uci2017} and the Cleveland Heart Disease
dataset~\cite{detrano1989}.
Experimental code (including all five training configurations, tuning scripts,
and benchmark runner) is available at
\url{https://github.com/CarlosMelroLeandro/luknn-extended}.

% ── Bibliography ──────────────────────────────────────────────────────────────
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
